---
title: "Lab 6 - Alternative"
subtitle: "Contributing to R for Data Science"
auther: Lucas Kantorowski
editor: source
error: true
warning: false
eval: false
---

The exercises from the functions below were pulled from the newest version of
[*R for Data Science*](https://r4ds.hadley.nz/). Specifically, from Chapters 25 
and 26. For this "alternative" lab you will complete the exercises from the 
textbook, with the option of submitting a pull request to the 
[repository for the textbook solutions](https://github.com/mine-cetinkaya-rundel/r4ds-solutions/).

```{r}
#| label: load_libraries

library(dplyr)
library('e1071')
library(tidyverse)
library(rlang)
```

# Vector Functions

**Question 1:** The `rescale01()` function below performs a min-max scaling to 
standardize a numeric vector, but infinite values are left unchanged. Rewrite
`rescale01()` so that `-Inf` is mapped to 0, and `Inf` is mapped to 1?
*Hint: This seems like a great place for `case_when()`!

```{r}
#| label: question-1


rescale01 <- function(x) {
  if (class(x) != "numeric") {
    stop("The provided vector is not numeric")
  }
  
  if (length(x) < 2) {
    stop("The provided vector is less than length 2")
  }
  
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  
  #making negative infinite and infinite values the min and max, respectively
  x = case_when(x == -Inf ~ rng[1], x == Inf ~ rng[2],.default = x)
  
  (x - rng[1]) / (rng[2] - rng[1])
  
  
}
```

```{r}
#testing with NA values
rescale01(c(1,2,3,4,5,6,NA,-7,8,Inf, NA))

#testing with negative infinity and negative inifinity values
rescale01(c(1,2,3,-Inf,5,6,NA,-7,8,Inf, NA))

#testing with character values
rescale01(c(1,2,3,-Inf,5,'6',NA,-7,8,Inf, NA))

#tseting with vector of length 1
rescale01(c(1))

```

**Question 2:** Write a function that accepts a vector of birthdates and 
computes the age of each person in years.

```{r}
#| label: question-2

age_in_years = function(date_vect) {
  return(date_vect)
  
  #calculate year difference, then round down
  
}

apple = c("01/01/2000", "10/23/2001")
class(apple)
apple

banana = as.Date(apple, format = "%m/%d/%yy")
banana

class(apple)
class(banana)
age_in_years(c(banana))
year(banana)

date()
difftime(date(), date())
```

**Question 3:** Write a function that computes the variance and skewness of a
numeric vector. Feel free to look up the definitions on Wikipedia or elsewhere!

```{r}
#| label: question-3

variance_and_skewness = function(num_vec){
  
  if (class(num_vec) != "numeric") {
    stop("The provided vector is not numeric")
  }
  
  
  v_s_df = data.frame(Variance = var(num_vec, na.rm = TRUE), 
                      Skewness = skewness(num_vec, na.rm = TRUE))
  
  #return(var(num_vec), skewness(num_vec))
  return(v_s_df)
}
```

```{r}

#testing with normal vector
variance_and_skewness(c(1,2,3,4,5,7,8,89,9,87,0))

#testing with a character vector
variance_and_skewness(c(1,2,3,4,5,7,8,89,9,87,'0'))

#testing with a vector of length 1
variance_and_skewness(c(1))

#testing with NA values
variance_and_skewness(c(1,2,3,4,5,NA,8,89,9,87))

```

**Question 4:** Write a function called `both_na()` which takes two vectors of
the same length and returns the number of positions that have an `NA` in both
vectors.


```{r}
both_na = function(vec1, vec2){
  
  #checking if columns are same length
  if (length(vec1) != length(vec2)) {
    stop("The provided vectors are not the same length")
  }
  
  
  #check for if a vector is non-numeric
  if (class(vec1) != "numeric" | (class(vec2) != "numeric"))  {
    stop("At least 1 provided vector is not numeric")
  }
  
  
  #adding up the NA status of column 1 and column 2, then filtering by col3 = 2
  return(data.frame(vec1, vec1, 
             row_num_nas = (is.na(vec1) + is.na(vec2))) %>% 
    filter(row_num_nas == 2) %>% #  filtering by is.na(col1) + is.na(col2) = 2
    summarise(N = n()) %>% as.numeric()) # counting number of columns with 2 NAs
  
  
}
#extract single value from 1x1 data frame: https://stackoverflow.com/questions/60864283/extract-the-single-value-from-a-1-x-1-data-frame-produced-with-dplyr-as-a-vector


```

```{r}

#testing with basic input
vec_a = c(NA,3,4,6, NA, NA, NA, 11, NA)
vec_b = c(NA, NA, NA,3,4, 5,NA,3,5)
both_na(vec_a, vec_b)


#testing with 0 values
vec_a = c(1,2,3,4,5,6)
vec_b = c(64, 78, 456456, 88, 44, 29)
both_na(vec_a, vec_b)


#testing with 1 vector having character values
vec_a = c(NA,NA,3,4,5,6)
vec_b = c(NA, NA, NA, '2',3,4)
both_na(vec_a, vec_b)


#testing with different length vectors
vec_a = c(NA,3,4,5,6)
vec_b = c(NA, NA, NA, '2',3,4)
both_na(vec_a, vec_b)

```



## Data Frame Functions

**Question 5:** Insert the data frame function you wrote from Lab 6 (either
Exercise 1 or Exercise 2). 

```{r}
#| label: function-from-lab-2

impute_missing = function(df, ..., impute_fun = mean) {
  # Check to verify function input is a function
  if (!is.function(impute_fun)) {
    stop("The input for `impute_fun` must be a function.")
  }
  
  # Check for non-character columns (from live code functions.qmd)
  character <- df %>%
    select(...) %>%
    map_lgl(~ (is.character(.x) | is.factor(.x))) %>%
    keep(identity) %>%
    names()
  
  if (length(character) > 0) {
    stop("The following columns are characters: ",
         str_c(character, collapse = ", "))
  }

  newdf = df %>% 
    mutate(across(c(...), ~ replace_na(data = .x, 
                                       replace = impute_fun(.x, na.rm = TRUE))))
  return(newdf)

}
```

```{r}

#testing with 1 numeric column
mean(nycflights13::flights$arr_delay, na.rm = TRUE)

test1 = impute_missing(nycflights13::flights, 
               arr_delay)

test1 %>% select(arr_delay) %>% group_by(arr_delay) %>% 
  summarise(N = n()) %>% arrange(desc(N)) %>% slice_head(n = 10)


#testing with non-default function
median(nycflights13::flights$arr_delay, na.rm = TRUE)

test2 = impute_missing(nycflights13::flights, 
               arr_delay, impute_fun = median)

test2 %>% select(arr_delay) %>% group_by(arr_delay) %>% 
  summarise(N = n()) %>% arrange(desc(N)) %>% slice_head(n = 10)


#testing with 2 numeric columns
test3 = impute_missing(nycflights13::flights, 
               arr_delay, dep_delay)

test3 %>% select(arr_delay) %>% group_by(arr_delay) %>% 
  summarise(N = n()) %>% arrange(desc(N)) %>% slice_head(n = 10)

test3 %>% select(dep_delay) %>% group_by(dep_delay) %>% 
  summarise(N = n()) %>% arrange(desc(N)) %>% slice_head(n = 10)


#testing with character column input
impute_missing(nycflights13::flights,carrier)

impute_missing(nycflights13::flights, 
               arr_delay, dep_delay, carrier)


#testing with non-real impute function
impute_missing(nycflights13::flights, dep_delay, impute_fun = apple)

```

For Questions 6 - 10 you will write different functions which work with data 
similar to the `nycflights13` data. 

**Question 6:** Write a `filter_severe()` function that finds all flights that
were cancelled (i.e. `is.na(arr_time)`) or delayed by more than an hour.

```{r}
#| label: question-6

filter_severe = function(df = nycflights13::flights){
  
  
  if ('dep_delay' %in% colnames(df) == FALSE | 
      'arr_time' %in% colnames(df) == FALSE){
    stop("Dataset provided does not have desired column names")
  }
  
  return(df %>% filter(is.na(arr_time) == TRUE | dep_delay > 60))
}
```

```{r}

#testing for normal inputs
filter_severe(nycflights13::flights)


#testing for unexpected dataframe
filter_severe(mpg)

```

**Question 7:** Write a `summarize_severe()` function that counts the number of 
cancelled flights and the number of flights delayed by more than an hour.

```{r}
#| label: question-7

summarize_severe = function(df = nycflights13::flights){
  #using previous function
  #counting number of cancelled flights
  #counting number of delayed flights by more than one hour
  
  return(filter_severe(df) %>% 
           summarise(Cancelled = sum(is.na(arr_time)),
                                         Delayed = sum(dep_time > 60, 
                                                       na.rm = TRUE)))
}
```

```{r}

#testing expected input
summarize_severe(nycflights13::flights)


#testing function without necessary columns
summarize_severe(mpg)

```

**Question 8:** Modify your `filter_severe()` function to allow the user to 
supply the number of hours that should be used to filter the flights that were
cancelled or delayed. 

```{r}
#| label: question-8

filter_severe = function(df = nycflights13::flights, hours = 1){
  if ('dep_delay' %in% colnames(df) == FALSE | 
      'arr_time' %in% colnames(df) == FALSE){
    stop("Dataset provided does not have desired column names")
  }
  
  if (class(hours) != 'numeric'){
    stop("Please provide a numeric qunaitity for delay hours")
  }
  
  
  return(df %>% filter(is.na(arr_time) == TRUE | dep_delay > hours * 60))
  
  
}
```

```{r}
#testing for normal inputs
filter_severe(nycflights13::flights, 1)
filter_severe(nycflights13::flights, 10)

#testing for non-numeric delay
filter_severe(nycflights13::flights, '1')


#testing for unexpected dataframe
filter_severe(mpg, 4)

```

**Question 9:** Write a `summarize_weather()` function that summarizes the
weather to compute the minimum, mean, and maximum, of a user supplied variable. 

```{r}
#| label: question-9

summarize_weather = function(df = nycflights13::flights, num_var){
  #make sure variable is numeric
  
  if (length(df %>% select({{num_var}}) %>% 
             purrr::discard(~!is.numeric(.))) == 0){
    stop('Please provide a numeric variable')
  }
  

  
  return(df %>% select({{num_var}}) %>%
            summarise(Minimum = min({{num_var}}, na.rm = TRUE),
            Mean = mean({{num_var}}, na.rm = TRUE),
            Maximum = max({{num_var}}, na.rm = TRUE)))
}  
```

```{r}
  
#testing with normal input
summarize_weather(nycflights13::flights, dep_time)


#testing with non-numeric column
summarize_weather(nycflights13::flights, carrier)
```

**Question 10:** Write a `standardize_time()` function that converts the user
supplied variable that uses clock time (e.g., `dep_time`, `arr_time`, etc.) into
a decimal time (i.e. hours + (minutes / 60)).

```{r}
#| label: question-10

standardize_time = function(df = nycflights13::flights, time_var){
  #user supplied variable that uses clock time into decimal time
  if (englue('{{time_var}}') %in% 
      c('dep_time', 'sched_dep_time', 'arr_time', 'sched_arr_time') == FALSE){
    stop("Please provide a clock time variable to summarize")
  }
  
  return(df %>% 
    mutate({{time_var}} := floor({{time_var}}/100) * 60 + {{time_var}} %% 100))
}
```


```{r}
#testing using expected input
standardize_time(nycflights13::flights, arr_time)
standardize_time(nycflights13::flights, sched_dep_time)

#testing supplying a non-time column
standardize_time(nycflights13::flights, day)
```

# Plotting Functions

You might want to read over the [Plot Functions section of *R for Data Science*](https://r4ds.hadley.nz/functions.html#plot-functions)

**Question 11:** Build a `sorted_bars()` function which:

- takes a data frame and a variable as inputs and returns a **vertical** bar
chart 
- sorts the bars in decreasing order (largest to smallest)
- adds a title that includes the context of the variable being plotted

*Hint 1: The `fct_infreq()` and `fct_rev()` functions from the forcats package will be helpful for sorting the bars!*
*Hint 2: The `englue()` function from the rlang package will be helpful for adding a variable's name into the plot title!*

```{r}
#| label: question-11

sorted_bars <- function(df, var) {
  
  #checking to make sure input variable is categorical
  if (length(df %>% select({{var}}) %>% purrr::discard(~is.numeric(.))) == 0){
    stop('Please provide a categorical variable')
  }
  
  #creating name object for label purposes
  name = englue("{{ var }}")
  
  return(df %>% select({{var}}) %>% mutate({{var}} := fct_infreq({{var}})) %>% 
    ggplot(aes(x = {{var}})) + 
    ggtitle(paste("Distribution of", name, "(decreasing)")) + 
    geom_bar())
  
}  
```

```{r}
#keeping only numeric variables
#https://stackoverflow.com/questions/5863097/selecting-only-numeric-columns-from-a-data-frame

#testing using a expected input
sorted_bars(nycflights13::flights, carrier)
sorted_bars(mpg, trans)


#testing using a numeric variable
sorted_bars(nycflights13::flights, dep_delay)

```

# Iteration

Alright, now let's take our plotting function and iterate it! 

**Question 12:** Make a sorted barplot for every character variable in the `mpg`
dataset (built into `ggplot2`). 

```{r}
#| label: question-12

#could not get sorting to work with map function
unsorted_bars <- function(df, var) {
  
  #checking to make sure input variable is categorical
  if (length(df %>% select({{var}}) %>% purrr::discard(~is.numeric(.))) == 0){
    stop('Please provide a categorical variable')
  }
  
  #creating name object for label purposes
  name = englue("{{ var }}")
  
  return(df %>% select({{var}}) %>% 
    ggplot(aes(x = {{var}})) + 
    ggtitle(paste("Distribution of", name, "(decreasing)")) + 
    geom_bar())
  
}  
```


```{r}
#barplots on all categorical variables of mpg
mpg %>% purrr::discard(~is.numeric(.)) %>% names() %>% map(~unsorted_bars(df = mpg, var = .data[[.x]]))

#use of .data[[.x]]
#https://stackoverflow.com/questions/60305159/tunnel-data-variables-through-env-variable-with-operator

```



# Contributing to the R for Data Science Community!

The functions you wrote for exercises 1-10 came from *R for Data Science*. You
could consider making a pull request to the repository for the solutions! 

<https://github.com/mine-cetinkaya-rundel/r4ds-solutions>

To learn more about how to make a pull request I would suggest this article: <https://usethis.r-lib.org/articles/pr-functions.html>
